Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations

### 2016-03-16 doc/manuscript.tex
move figures
add caption for new figure 1

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations},
            pdfauthor={Thomas McColgan, Hermann Wagner, Ji Liu, Paula T Kuokkanen, Catherine E Carr, Richard Kempter},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing

\title{Extracellular potentials generated by axonal projections are shaped by
patterns of bifurcations and terminations}
\author{Thomas McColgan, Hermann Wagner, Ji Liu, Paula T Kuokkanen, Catherine E
Carr, Richard Kempter}
\date{\today}

@@

The aim of this study is to show how dipolar far fields in the EFP of
axons can be explained by the axons' anatomical structure. In
particular, we show that the branchings and termonations of axons in
their projection area lead to a dipolar EFP structure. Such axon
bundles, often also called nerves or fascicles, exist throughout the
peripheral and central nervous system of most animals (Nornes and Das,
1972; Goodman et al., 1984; Hentschel and Ooyen, 1999; Kandel et al.,
2000). The white matter of the mammalian brain can be viewed as an
agglomeration of such bundles (Schüz and Braitenberg, 2002). We
therefore predict pronounced contributions of nerves or fascicles to
EFPs, which are neglected in all current models.

@@

\emph{Axonal projections generate a dipole-like field potential.} To
understand how axons can generate a dipolar EFP, we simulated a parallel
fiber bundle that at first runs at constant diameter without
bifurcations and then reaches a projection zone. Within this projection
zone, the fibers first bifurcate, which increases the number of fibers.
Finally, the axons reach a termination zone, in which the number of
fibers decreases to zero (Figure 1A).

@@

To characterize the spatiotemporal dynamics of the evoked EFP, the
potential was calculated for several locations along the axon trunk.
Because we first want to study the effect of the rate pulse on the EFP,
we low-pass filtered the EFP responses with a cut-off frequency of 1 kHz
(Figure 1B). The distribution of maximum amplitudes of these responses
are shown by the colored contour lines in Figure 1A and the colored
traces in Figure 1B. Surrounding the projection zone of the axon bundle,
amplitudes show the double-lobed shape typical of a dipole.

In Figure 1B, all the responses showed a biphasic deflection that was
elicited by the population firing rate pulse. This deflection was
visible at all recording locations. The biphasic deflection reverses
polarity in the middle of the bifurcation zone of the bundle (red box in
Figure 1B). The polarity reversal occurs by a decrease of the amplitude
to zero and subsequent reappearance with reversed polarity, as opposed
to a polarity reversal through a gradual shift in phase. This behaviour
is also typical for a dipolar field potential.

The point of the polarity reversal coincides roughly with the middle of
the projection zone. This means that the absolute value of the response
amplitude reaches a local minimum just when the number of axonal fibers
reaches a maximum. To better understand how anatomical features and
response amplitude are related, we compared the signed maximum value
(meaning the \emph{signed} value corresponding to the maximum
\emph{absolute} amplitude) of the EFP with the difference in the density
of branchings and terminations (Figure 1C).

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_1.pdf}
\caption{Axonal projections generate a dipole-like extracellular field
potential (EFP) when stimulated with a pulse of activity. \textbf{A}
Modeled axon bundle in black, along with iso-amplitude lines for the
low-pass filtered EFP signature of the activity pulse. The contours
(maximum amplitude in mV as indicated by colorbar) show the typical
double-lobe of a dipole. \textbf{B} The waveforms, recorded at the
locations of the colored dots in \emph{A} also show a polarity reversal
(red box). The reversal ocurrs by inverting the envelope with
aproximately constant phase, not by a shifting of phase with constant
envelope. \textbf{C} Progression of the maximum amplitude with depth
(red line). It closely follows the local difference in density between
bifurcations and terminations (black histogram). \textbf{D} Iso-amlitude
contours for the high-frequency component, which do not exhibit a
doiuble-lobed structure. \textbf{E} Response waveforms of the
low-frequency component. The responses do not reverse their polarity.
\textbf{F} Maximum amplitude of the high-frequency component (red line)
follows the density of fibers (black histogram).}
\end{figure}

To understand the contributions related to individual spikes, we next
turned our attention to the high-frequency range. We filtered the EFP
responses with a high-pass filter, using the same cut-off frequency of
1kHz as before. The maximum amplitudes for this frequency band did not
show a double-lobe, but rather appeared like an ellipsoid centered at
the projection zone (Figure 1D). The waveforms are more disparate across
recording locations, suggesting a more local nature. There is no visible
polarity reversal (Figure 1E).

The high-frequency component changes its amplitude in the longitudinal
direction in accordance to the local fiber density (Figure 1F). It has
an approximately constant amplitude along the nerve trunk and then
increases in amplitude as the fiber density is increased by
bifurcations. As the fibers terminate and fiber density decreases, so
does the amplitude of the high-frequency component.

We have thus shown a qualitatively different behaviour in the low- and
high-frequency components of the EFP. Both appear to be related to
anatomy, though each in a different way. Together, these relationships
suggest that the particular branching and terminating structure of the
axon bundle gives rise to the dipolar field potential. Because dipolar
fields are associated with a longer spatial reach than the quadrupolar
fields typically associated with axonal spikes, we decided to
investigate the effect that axon structure has on the spatial reach of
the frequency components (Figure 2). To do this we performed a further
simulation of an axon bundle, this time omitting the bifurcations, and
letting the fibers terminate directly. In axial direction, the
low-frequency component has a further reach than the high frequency
component in the bifurcating case (Figure 2A), but not in the
non-bifurcating case (Figure 2C). In the radial direction, the situation
is different: the bifurcating (Figure 2B) and non-bifurcating (Figure
2D) case both show a faster decay of the high-frequency component than
of the low-frequency component. To assure that the effect was due to the
bifurcations themselves and not due to the fanned out structure of the
bifurcating bundle seen in Figures 2A and 2D, we here changed the layout
of both bundles to have constant widths.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_2.pdf}
\caption{Low-frequency component of the axon bundle EFP exceeds
high-frequency component in reach, modulated by axon structure. All
plots show the scaling of the low-frequency (blue lines) and
high-frequency (green lines) frequency components on a
doubly-logarithmic scale. Amplitudes were normalized to unit value at
the end of the tuft in axial direction (\textbf{A} and \textbf{B}) and
from the trunk in radial direction (\textbf{C} and \textbf{D}).
\textbf{A} and \textbf{C} show the behaviour in the case of a
bifurcating bundle, while \textbf{B} and \textbf{D} show the
nonbifurcating case. In the bifurcating case, the low-frequency
component exceeds the reach of the high-frequency comonent in both
directions. In the non-bifurcating case, the decay is the same for both
components in the axial direction. The scaling behaviours are consistent
with the presence of a low-frequency dipole-like potential in the
bifurcating bundle but not in the non-bifurcating bundle.}
\end{figure}

@@

We adjusted the model parameters to match the anatomy and physiology of
nucleus laminaris (NL), and set the activation statistics to match
measured PSTHs in response to stimulation by a click sound presented to
one ear. We then simulated the resulting EFP at varying depths,
corresponding to recording depths in the experiment. The resulting
recordings from experiment and simulation are shown in Figure 3.

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_3.pdf}
\caption{Data from the barn owl shows the expected behaviour predicted
by the model. (\textbf{A},\textbf{B}) shows data from the barn owls
nucleus laminaris in response to an auditory click stimulus, compared to
a simulation of the axonal structure and activation in
(\textbf{C},\textbf{D}). The click stimulus induces a pulse of activity
in the afferent axon bundle. The low-frequency components (\textbf{A}
and \textbf{C}) show the polarity reversal. The high frequency component
(\textbf{B} and \textbf{D}), does not show such a reversal, but rather
shows a steady increase in phase with depth.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/mockups/fig4.pdf}
\caption{The dipolar behaviour can be understood by examining individual
action potentials on a single axon tree. Comparing the low frequency owl
data (\textbf{A}) to a single axon and action potential in model
(\textbf{B}) shows a similar behaviour. In particular, the potential at
a termination and that at a bifurcation (red and green curves in
\textbf{B}) are approximately inverted.}
\end{figure}

@@

Further inspection of the two components shows an intriguing similarity
to the properties of the response observed in the previous sections. The
first component is high-pass filtered, and is related to the fiber
density, just like `noise' component in Figure 1 and the high-frequency
component in the barn owl data and model. The second component is
low-pass filtered, and related to the derivative of the fiber density,
just like the firing rate pulse related deflection in Figure 1, and the
low-frequency component in the barn owl.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/mockups/fig5.pdf}
\caption{Analytical model of the axon bundle potential explains the
effects observed in the numerical model and example data. (\textbf{A})
shows the behaviour of a simplified fiber bundle with a piecewise
constant fiber density (\textbf{Aa}). The high-frequency component
(\textbf{Ab}) shows no polarity reversal, while the low-frequency
component (\textbf{Ac}) does, as expected from the data and modeling.
This can be understood by decomposing the signal into two components.
The first component is governed by the bifurcation and termination
density, and is filtered by the regular weighting function
(\textbf{Ba}), which acts as a low-pass filter (\textbf{Bb}). The second
component is governed by the fiber density, and is filtered by the
derivative of the weighting function (\textbf{Bc}), which acts as a
high- or band-pass filter (\textbf{Bd}).}
\end{figure}

@@

  \begin{itemize}
  \tightlist
  \item
    implement new simplification
  \item
    adapt generic bundle params to be more typical
  \item
    make click psth more generic
  \end{itemize}
\item
  Introduction

@@

  \begin{itemize}
  \tightlist
  \item
    Page 4 changes
  \item
    much more substantial explanation of Fig 2
  \item
    describe differences between barn owl and generic model
  \item
    Add more biological motivation for figure 2 normalization
  \end{itemize}
\item
  Figures

  \begin{itemize}
  \tightlist
  \item
    New Figure 1

      \begin{itemize}
      \tightlist
      \item
        straight
      \item
        branching
      \item
        terminating
      \end{itemize}
    \item
      compound
    \item
      CSD
    \item
      PSTH
    \end{itemize}
  \item
    Figure 1

    \begin{itemize}
    \tightlist
    \item
      version with MUA
    \item
      lower density of axons
    \item
      Add scalebar for time
    \item
      change color of axis label
    \item
      change color of box
    \item
      add PSTH
    \item
      align zeros in C
    \item
      add contour labels?
    \end{itemize}
  \item
    Figure 2

    \begin{itemize}
    \tightlist
    \item
      add diagrams
    \item
      space between B and C
    \item
      better location for axial scaling
    \end{itemize}
  \item
    Figure 5Bb log scale
  \end{itemize}
\item
  Methods

@@

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations},
            pdfauthor={Thomas McColgan; Ji Liu; Paula T Kuokkanen; Catherine E Carr; Hermann Wagner; Richard Kempter},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing
\usepackage{subfig}
\AtBeginDocument{%
\renewcommand*\figurename{Figure}
\renewcommand*\tablename{Table}
}
\AtBeginDocument{%
\renewcommand*\listfigurename{List of Figures}
\renewcommand*\listtablename{List of Tables}
}
\usepackage{float}
\floatstyle{ruled}
\makeatletter
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\makeatother
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}

\title{Extracellular potentials generated by axonal projections are shaped by
patterns of bifurcations and terminations}
\author{Thomas McColgan \and Ji Liu \and Paula T Kuokkanen \and Catherine E Carr \and Hermann Wagner \and Richard Kempter}
\date{\today}

@@

The aim of this study is to show how dipolar far fields in the EFP of
axons can be explained by the axons' anatomical structure. In
particular, we show that the branchings and terminations of axons in
their projection area lead to a dipolar EFP structure. Such axon
bundles, often also called nerves or fascicles, exist throughout the
peripheral and central nervous system of most animals (Nornes and Das,
1972; Goodman et al., 1984; Hentschel and Ooyen, 1999; Kandel et al.,
2000). The white matter of the mammalian brain can be viewed as an
agglomeration of such bundles (Schüz and Braitenberg, 2002). We
therefore predict pronounced contributions of nerves or fascicles to
EFPs, which are neglected in all current models.

@@

\emph{Axonal projections generate a dipole-like field potential.} To
understand how axons can generate a dipolar EFP, we simulated a parallel
fiber bundle that at first runs at constant diameter without
bifurcations and then reaches a projection zone. Within this projection
zone, the fibers first bifurcate, which increases the number of fibers.
Finally, the axons reach a termination zone, in which the number of
fibers decreases to zero (Figure~\ref{fig:bigtree}A).

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_1.pdf}
\caption{\label{fig:simpletree}Relationship between axon structure and
extracellular potential. The top row shows simulated action potentials
travelling along a single axon with varying morphologies, as indicated
by the diagram in the left side of each subfigure. Action potential
propagation direction indicated by the arrow. All waveforms recorded at
a distance of 150\(\mu\)m from the axon, and at the depth indicated by
plot position, spaced by 200\(\mu\)m. (\textbf{A}) An action potential
in an infinitely long, straight axon. (\textbf{B}) Effect of a
terminatiing axon. Action potential waveform closest to the termination
thickened for emphasis. (\textbf{C}) Effect of a branching axon. The
axon branches multiple times in the range of 200\(\mu\)m. Thicker
waveform at the center of the bifurcation zone. \textbf{D} Combined
effect of bifurcations and terminations. Lower row shows responses of a
population of 100 randomly branching and termnating axons. Three sample
morphologies are shown to the left. \textbf{E} Extracellular waveform in
response a spike initiated simultaneously in all axons. Spike initiation
time indicated by the raster lines at the top. \textbf{F} CSD calculated
from the waveforms shown in E. \textbf{G} Response to spontaneous
activity superimposed with a pulse of hightened spike rate, averaged
over 100 trials. Example spike times shown by the raster plot at
top.\textbf{H} CSD calculated from the pulse response in G.}
\end{figure}

To characterize the spatiotemporal dynamics of the evoked EFP, the
potential was calculated for several locations along the axon trunk.
Because we first want to study the effect of the rate pulse on the EFP,
we low-pass filtered the EFP responses with a cut-off frequency of 1 kHz
(Figure~\ref{fig:bigtree}B). The distribution of maximum amplitudes of
these responses are shown by the colored contour lines in
Figure~\ref{fig:bigtree}A and the colored traces in
Figure~\ref{fig:bigtree}B. Surrounding the projection zone of the axon
bundle, amplitudes show the double-lobed shape typical of a dipole.

In Figure~\ref{fig:bigtree}B, all the responses showed a biphasic
deflection that was elicited by the population firing rate pulse. This
deflection was visible at all recording locations. The biphasic
deflection reverses polarity in the middle of the bifurcation zone of
the bundle (red box in Figure~\ref{fig:bigtree}B). The polarity reversal
occurs by a decrease of the amplitude to zero and subsequent
reappearance with reversed polarity, as opposed to a polarity reversal
through a gradual shift in phase. This behaviour is also typical for a
dipolar field potential.

The point of the polarity reversal coincides roughly with the middle of
the projection zone. This means that the absolute value of the response
amplitude reaches a local minimum just when the number of axonal fibers
reaches a maximum. To better understand how anatomical features and
response amplitude are related, we compared the signed maximum value
(meaning the \emph{signed} value corresponding to the maximum
\emph{absolute} amplitude) of the EFP with the difference in the density
of branchings and terminations (Figure~\ref{fig:bigtree}C).

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_2.pdf}
\caption{\label{fig:bigtree}Axonal projections generate a dipole-like
extracellular field potential (EFP) when stimulated with a pulse of
activity. \textbf{A} Modeled axon bundle in black, along with
iso-amplitude lines for the low-pass filtered EFP signature of the
activity pulse. The contours (maximum amplitude in mV as indicated by
colorbar) show the typical double-lobe of a dipole. \textbf{B} The
waveforms, recorded at the locations of the colored dots in \emph{A}
also show a polarity reversal (red box). The reversal ocurrs by
inverting the envelope with aproximately constant phase, not by a
shifting of phase with constant envelope. \textbf{C} Progression of the
maximum amplitude with depth (red line). It closely follows the local
difference in density between bifurcations and terminations (black
histogram). \textbf{D} Iso-amlitude contours for the high-frequency
component, which do not exhibit a doiuble-lobed structure. \textbf{E}
Response waveforms of the low-frequency component. The responses do not
reverse their polarity. \textbf{F} Maximum amplitude of the
high-frequency component (red line) follows the density of fibers (black
histogram).}
\end{figure}

To understand the contributions related to individual spikes, we next
turned our attention to the high-frequency range. We filtered the EFP
responses with a high-pass filter, using the same cut-off frequency of
1kHz as before. The maximum amplitudes for this frequency band did not
show a double-lobe, but rather appeared like an ellipsoid centered at
the projection zone (Figure~\ref{fig:bigtree}D). The waveforms are more
disparate across recording locations, suggesting a more local nature.
There is no visible polarity reversal (Figure~\ref{fig:bigtree}E).

The high-frequency component changes its amplitude in the longitudinal
direction in accordance to the local fiber density
(Figure~\ref{fig:bigtree}F). It has an approximately constant amplitude
along the nerve trunk and then increases in amplitude as the fiber
density is increased by bifurcations. As the fibers terminate and fiber
density decreases, so does the amplitude of the high-frequency
component.

We have thus shown a qualitatively different behaviour in the low- and
high-frequency components of the EFP. Both appear to be related to
anatomy, though each in a different way. Together, these relationships
suggest that the particular branching and terminating structure of the
axon bundle gives rise to the dipolar field potential. Because dipolar
fields are associated with a longer spatial reach than the quadrupolar
fields typically associated with axonal spikes, we decided to
investigate the effect that axon structure has on the spatial reach of
the frequency components (Figure~\ref{fig:distscaling}). To do this we
performed a further simulation of an axon bundle, this time omitting the
bifurcations, and letting the fibers terminate directly. In axial
direction, the low-frequency component has a further reach than the high
frequency component in the bifurcating case
(Figure~\ref{fig:distscaling}A), but not in the non-bifurcating case
(Figure~\ref{fig:distscaling}C). In the radial direction, the situation
is different: the bifurcating (Figure~\ref{fig:distscaling}B) and
non-bifurcating (Figure~\ref{fig:distscaling}D) case both show a faster
decay of the high-frequency component than of the low-frequency
component. To assure that the effect was due to the bifurcations
themselves and not due to the fanned out structure of the bifurcating
bundle seen in Figure~\ref{fig:distscaling}A and
Figure~\ref{fig:distscaling}D, we here changed the layout of both
bundles to have constant widths.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_3.pdf}
\caption{\label{fig:distscaling}Low-frequency component of the axon
bundle EFP exceeds high-frequency component in reach, modulated by axon
structure. All plots show the scaling of the low-frequency (blue lines)
and high-frequency (green lines) frequency components on a
doubly-logarithmic scale. Amplitudes were normalized to unit value at
the end of the tuft in axial direction (\textbf{A} and \textbf{B}) and
from the trunk in radial direction (\textbf{C} and \textbf{D}).
\textbf{A} and \textbf{C} show the behaviour in the case of a
bifurcating bundle, while \textbf{B} and \textbf{D} show the
nonbifurcating case. In the bifurcating case, the low-frequency
component exceeds the reach of the high-frequency comonent in both
directions. In the non-bifurcating case, the decay is the same for both
components in the axial direction. The scaling behaviours are consistent
with the presence of a low-frequency dipole-like potential in the
bifurcating bundle but not in the non-bifurcating bundle.}
\end{figure}

@@

We adjusted the model parameters to match the anatomy and physiology of
nucleus laminaris (NL), and set the activation statistics to match
measured PSTHs in response to stimulation by a click sound presented to
one ear. We then simulated the resulting EFP at varying depths,
corresponding to recording depths in the experiment. The resulting
recordings from experiment and simulation are shown in
Figure~\ref{fig:barnowl}.

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_4.pdf}
\caption{\label{fig:barnowl}Data from the barn owl shows the expected
behaviour predicted by the model. (\textbf{A},\textbf{B}) shows data
from the barn owls nucleus laminaris in response to an auditory click
stimulus, compared to a simulation of the axonal structure and
activation in (\textbf{C},\textbf{D}). The click stimulus induces a
pulse of activity in the afferent axon bundle. The low-frequency
components (\textbf{A} and \textbf{C}) show the polarity reversal. The
high frequency component (\textbf{B} and \textbf{D}), does not show such
a reversal, but rather shows a steady increase in phase with depth.}
\end{figure}

@@

@@

Further inspection of the two components shows an intriguing similarity
to the properties of the response observed in the previous sections. The
first component is high-pass filtered, and is related to the fiber
density, just like `noise' component in Figure~\ref{fig:bigtree} and the
high-frequency component in the barn owl data and model. The second
component is low-pass filtered, and related to the derivative of the
fiber density, just like the firing rate pulse related deflection in
Figure~\ref{fig:bigtree}, and the low-frequency component in the barn
owl.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/mockups/fig5.pdf}
\caption{\label{fig:anamodel}Analytical model of the axon bundle
potential explains the effects observed in the numerical model and
example data. (\textbf{A}) shows the behaviour of a simplified fiber
bundle with a piecewise constant fiber density (\textbf{Aa}). The
high-frequency component (\textbf{Ab}) shows no polarity reversal, while
the low-frequency component (\textbf{Ac}) does, as expected from the
data and modeling. This can be understood by decomposing the signal into
two components. The first component is governed by the bifurcation and
termination density, and is filtered by the regular weighting function
(\textbf{Ba}), which acts as a low-pass filter (\textbf{Bb}). The second
component is governed by the fiber density, and is filtered by the
derivative of the weighting function (\textbf{Bc}), which acts as a
high- or band-pass filter (\textbf{Bd}).}
\end{figure}

@@

  \begin{itemize}
  \tightlist
  \item
    adapt generic bundle params to be more typical
  \item
    make click psth more generic
  \item
    check units
  \end{itemize}
\item
  Introduction

@@

  \begin{itemize}
  \tightlist
  \item
    move forward effect explanation and describe
    Figure~\ref{fig:simpletree}
  \item
    Page 4 changes
  \item
    much more substantial explanation of Figure~\ref{fig:distscaling}
  \item
    describe differences between barn owl and generic model
  \item
    Add more biological motivation for Figure~\ref{fig:distscaling}
    normalization
  \end{itemize}
\item
  Figures

  \begin{itemize}
  \tightlist
  \item
    Figure~\ref{fig:simpletree}

    \begin{itemize}
    \tightlist
    \item
      caption
    \item
      better alignment in D
    \end{itemize}
  \item
    Figure~\ref{fig:bigtree}

    \begin{itemize}
    \tightlist
    \item
      lower density of axons
    \item
      Add scalebar for time
    \item
      change color of axis label
    \item
      change color of box
    \item
      add PSTH
    \item
      align zeros in C
    \item
      add contour labels?
    \end{itemize}
  \item
    Figure~\ref{fig:distscaling}

    \begin{itemize}
    \tightlist
    \item
      add diagrams
    \item
      space between B and C
    \item
      better location for axial scaling
    \end{itemize}
  \item
    Figure~\ref{fig:anamodel}Bb log scale
  \end{itemize}
\item
  Methods

@@

\section{Done}\label{done}

  \begin{itemize}
  \tightlist
  \item
    pandoc figure refs
  \item
    implement new simplification
  \end{itemize}
\item
  Figures

  \begin{itemize}
  \tightlist
  \item
    New Figure~\ref{fig:bigtree}

      \begin{itemize}
      \tightlist
      \item
        straight
      \item
        branching
      \item
        terminating
      \end{itemize}
    \item
      compound
    \item
      single axons

      \begin{itemize}
      \tightlist
      \item
        multibranch
      \item
        endings and snips
      \end{itemize}
    \item
      Figures E and F
    \item
      PSTH
    \item
      CSD
    \end{itemize}
  \item
    Figure~\ref{fig:bigtree}

    \begin{itemize}
    \tightlist
    \item
      version with MUA
    \end{itemize}
  \end{itemize}
\end{itemize}
